---
title: "Encode point color from values in your data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Encode point color from values in your data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The rscatter function has an argument `color` to set the point color to a static color value.

```{r static, out.width="100%"}
library(rscatter)
data("faithful")
rscatter("waiting", "eruptions", data = faithful, color = "peachpuff3")
```

However often you will want to encode the point color based on the different values in your dataset, from both categorical and continuous variables.

To encode point color based on categorical values in your data make sure the variable is a character or factor.

```{r categorical, out.width="100%"}
data("quakes")
rscatter(quakes$long, quakes$lat, colorBy = cut(quakes$depth, 2))
```

The above plot categorizes the depth of the quake into two categories. To visualize the depth variable using a continuous color map leave it as a numeric variable.

```{r continuous, out.width="100%"}
rscatter(quakes$long, quakes$lat, colorBy = quakes$depth)
```

In both of these cases we didn't specify the exact colors to use, defaults were selected. However you can specify the colors you want.

For categorical color mapping pass a named color vector to the `color` argument of the `rscatter` function. Reusing our example from the introductory vignette we will explicitly set the color for the branches of the spiral.

```{r named, out.width="100%"}
phi = seq(from = 0, to = 50, by = 1e-2)
fermat_spiral = data.frame(
 x = c(-sqrt(phi)*cos(phi), sqrt(phi)*cos(phi)),
 y = c(-sqrt(phi)*sin(phi), sqrt(phi)*sin(phi)),
 branch = rep(c("a", "b"), each = length(phi))
 )
my_colors <- c("a" = "#FCFF8B", "b" = "#F2ACE0")
rscatter("x", "y", size = 1.5, color = my_colors, colorBy = "branch", data = fermat_spiral)
```

For a continuous color map send the palette to the `color` argument.

```{r colorRamp, out.width="100%"}
n = seq(from = 0, to = 2*pi, by = 0.1)
y = sin(n)
dist = abs(y)
YlOrBr <- c("#FFFFD4", "#FED98E", "#FE9929", "#D95F0E", "#993404")
my_colors <- colorRampPalette(YlOrBr, space = "Lab")(256)
rscatter(n, y, color = my_colors, colorBy = dist)
```

