---
title: "Getting started with rscatter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with rscatter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The `rscatter` package is an R package that creates an [htmlwidget](https://www.htmlwidgets.org/) wrapping the [regl-scatterplot](https://github.com/flekschas/regl-scatterplot) JavaScript library. An htmlwidget widget outputs HTML, CSS, and JavaScript for viewing in various contexts that support HTML such as the viewer pane in RStudio, interactive R Markdown and Quarto documents, and Shiny apps.

In the case of rscatter the API is similar to other plotting functions in R and HTML is rendered that creates pan-and-zoomable scatterplots that scale to millions of points.

## Installation

The rscatter package is not yet on CRAN. You may install it from the GitHub repository.

```{r install, eval=FALSE}
# Install remotes package if necessary
if (!require("remotes")) {
  install.packages("remotes")
}

remotes::install_github("davidpross/rscatter", upgrade = FALSE)
```

## Basic usage

You can view the full documentation for the rscatter function [here](https://davidpross.github.io/rscatter/reference/rscatter.html) but you can get going quickly by simply passing your x and y coordinates.

```{r basic, out.width="100%"}
library(rscatter)
rscatter(rnorm(1e4), rnorm(1e4))
```

You can interact with the rendered plot by zooming and panning it.

## Plotting values in a data.frame

Many times during exploratory data analysis we have our data in a data.frame object. You can plot the values from your data.frame as shown in the following example.

```{r spiral, out.width="100%"}
phi = seq(from = 0, to = 50, by = 1e-2)
fermat_spiral = data.frame(
 x = c(-sqrt(phi)*cos(phi), sqrt(phi)*cos(phi)),
 y = c(-sqrt(phi)*sin(phi), sqrt(phi)*sin(phi)),
 branch = rep(c("a", "b"), each = length(phi))
 )
rscatter("x", "y", size = 1.5, colorBy = "branch", data = fermat_spiral)
```

Even with ~10K points the plot renders quickly and is responsive when interacting with it. To learn about the `colorBy` parameter used here see [the next article](https://davidpross.github.io/rscatter/articles/color-encoding.html).


